Coalition Formation Algorithm of Prosumers in a Smart Grid Environment

### 2014-09-17 Article.tex
update Vincent + Monique

As visible on figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P^{MIN}$ could potentially lead to valid coalitions (and positive utility values). Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and valid ones can arise for a larger spectrum of $ P^{MIN}$. Obviously, the highest utility values are found for high $ \phi $, because coalitions are able then to announce higher contract values, yielding higher utilities.

In the following, we fix the reliability to a given empirical value ($\phi = Cste $) and observe how the coalitions evolves for different values of $P^{MIN}$. Note that the opposite is also possible although a little less intuitive, but not shown here because of space limitations and because it leads to similar conclusions.

On figure \ref{fig2:b}, we can see that when the grid constraints are fixed ($ \phi = 0.08 $ and $ P^{MIN} = 2 $ for instance) and only the number of coalitions varies, there is an increase in social welfare (sum of coalitions utilities) up to a maximum point (for $ N_{COAL} = 3 $ here) before a decrease. The reason is that increasing the number of coalitions allows more coalitions to reach stability and enter the market, but there is a point where nodes bringing stability are not sufficient inside the coalitions to make them pass the grid requirements, and some coalitions start to fail with zero utility. Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs where information is flooded, meaning that the algorithm performances also decrease.
Naturally, it is also visible that small values of $P^{MIN}$ lead to higher utilities because coalitions are able to announce higher contract values.

\begin{center}
\begin{figure}[b]
  \includegraphics[scale=0.8]{./figure5/coals}
  \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots). The diameters of the dots are proportional to the corresponding coalition sizes.}
  \label{Fig3}
\end{figure}
\end{center}

For our 200 prosumers, we suppose that the grid fixes a reliability $ \phi = 0.1 $ (meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time). We also fix the number of coalitions to 10 and have a look at how these coalitions evolve as the grid changes the $ P^{MIN} $ requirement. As a comparison, we use a completely random algorithm that only asks for a number of coalitions and partitions the agents in a random fashion. Figure \ref{Fig3} shows this evolution for our algorithm (blue dots) and for the random process (red dots). The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot is, the higher its utility. The $ P^{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the "under-requirement space", meaning that whenever a coalition is in this zone, it has a null utility.

Looking at figure \ref{Fig3}, we can see first that our algorithm seems to make generally a better job at finding high valued coalitions, but also that the results seem far more robust against how the grid positions its requirements. The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither too low or too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).

In more details, figure \ref{fig4:a} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve), clique percolation (blue curve), and correlated coalitions (green curves) that stands for a worst case scenario. As for figure \ref{fig4:b}, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $. For consistency, we average the results of both plots of the random procedure over 100 realizations, and the errorbars in the plots stands for the standard deviations of the results.

When the grid requirements are constant (figure \ref{fig4:b}), and the number of desired coalitions is low, clique percolation generally performs only a little better than random. Nevetheless, when $ N_{COAL} $ gets bigger, the performance of a random split tumble down rapidely while clique percolation social welfare decreases slowly.

When the grid requirements vary, for very low $ P^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $. But as $ P^{MIN} $ increases, we see the percentage of the correlated coalitions crashing down very quickly. After a few increases in $ P^{MIN} $, the percentage of the random procedure starts droping while it stays almost constant for our algorithm. For $ P^{MIN} = 8 $, we see that only a little more than half of the coalitions for the random case are able to enter while approximately $ 85 \%$ of them enters for the clique percolation algorithm. Finally, when the grid requirement becomes too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).

>>>>

As shown in figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P^{MIN}$. This could potentially lead to valid coalitions (and positive utility values). Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions can arise for a larger spectrum of $ P^{MIN}$. Obviously, the highest utility values are found for high $ \phi $, because the coalitions formed are able then to announce higher contract values, yielding higher utilities.

In figure \ref{fig2:b}, we fix the reliability value to a given empirical value ($\phi = 0.08 $). We can observe how the social welfare $\sum_{i \in S} \tau_i(S)$ evolves according to $\phi$ and $P^{MIN}$. In our setting we can observe  a maximum point for $ N_{COAL} = 3 $, followed a decrease. There is clearly a maximum number of coalitions sustainable for given set of parameters. After that point we will form more coalitions that would be unable to pass the grid requirements. Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs, meaning that the algorithm performance will also decrease.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.8]{./figure5/coals}
  \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots). The diameters of the dots are proportional to the corresponding coalition sizes.}
  \label{Fig3}
\end{figure}

In figure \ref{Fig3}, we perform a simulation with 200 prosumers, and we fix the reliability $ \phi = 0.1 $ meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time. We observe the result of our algorithm for a fixed number of coalitions to $N_{coal} = 10$ and let the algorithm form the coalitions as the grid changes according to $ P^{MIN} $. We compare our algorithm with a random process that forms coalition. Figure \ref{Fig3} shows this evolution for our algorithm (blue dots) and for the random process (red dots). The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot is, the higher its utility. The $ P^{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the "under-requirement space", meaning that whenever a coalition is in this zone, it has a null utility.

Looking at figure \ref{Fig3}, we can see first that our algorithm performs better at finding high valued coalitions. The blue dots allowed to enter the market are indeed outnumbering the red dots, especially when the grid requirements are neither too low nor too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).

In more details, figure \ref{fig4:a} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve), clique percolation (blue curve), and correlated coalitions (green curves) that stands for a worst case scenario. As for figure \ref{fig4:b}, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $. For consistency, we average the results of both plots of the random procedure over 100 realizations with 95\% of confidence interval.

When the grid requirements are constant (figure \ref{fig4:b}), and the number of desired coalitions is low, clique percolation generally performs better than random case. Moreover, when $ N_{COAL} $ gets bigger, the performance of a random split tumble down rapidly while our clique percolation is able to maintain efficiently the social welfare of the coalitions formed.

When the grid requirements vary, for very low $ P^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $. But as $ P^{MIN} $ increases, we see the percentage of the correlated coalitions is going down very quickly. After a few increases in $ P^{MIN} $, the percentage of the random procedure starts dropping while it stays almost constant for our algorithm. For $ P^{MIN} = 8 $, we see that only a little more than half of the coalitions for the random case are able to enter while approximately $ 85 \%$ of them enters for the clique percolation algorithm. Finally, when the grid requirements becomes too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).
